---
title: "Google Trends Analysis in R"
---

# Task One: Getting Started
In this task, you will import the required
packages for this project
```{r}

## 1.1: Importing required packages
library(gtrendsR)
library(dplyr)
library(ggplot2)
library(ggforce)
library(ggrepel)
library(spData) # to load the world data
library(tmap)   # to map visualization
library(DT)     # to create the data tables shown at the bottom chunk


```

#Task Two: Getting Google trends data
In this task, we will learn how to get 
Google trends data about Covid
```{r}

## 2.1: Retrieve the worldwide data for 
## "covid" in the last 12 months
trends <- gtrends(keyword = c("covid","vaccine"), time = "today 12-m", geo = "US")
wtrends <- gtrends(keyword = c("covid","vaccine"), time = "today 12-m")
## 2.2: Return the names of the datasets from trends
names(trends)

## 2.3: Search for the keywords "covid" and 
## "vaccine" in US in the last 12 months
gtrends(keyword = c("covid","vaccine"),time = "today 12-m", geo = "US")

```

#Task Three: Interest Over Time Data
In this task, we will learn how to examine 
and explore the interest over time data
```{r}

## 3.1: Return the interest over time 
## data from the trends dataset
interest_time <- trends$interest_over_time

## 3.2: Print the interest_time data
interest_time

## 3.3: Return the dates with the highest number of hits
interest_time %>%
  top_n(5,hits) %>%
  arrange(desc(hits))

mx = interest_time[interest_time$hits == max(interest_time$hits),]$date

```

#Task Four: Visualize the Interest Over Time Data
In this task, we will learn how to 
visualize the interest over time data
```{r}

## 4.1: Visualize the interest over time data
interest_time %>%
  ggplot(aes(x = date, y = hits, color = keyword)) +
  geom_smooth(span = 0.1, se = FALSE) +
  geom_point(color = "black") +
  geom_vline(xintercept = as.POSIXct(mx), color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Google Web search for 'Covid & Vaccine' in 2021",
       x = NULL, y = "Interest",
       caption = paste("Google Web Search Analysis :",Sys.Date()))
 

  
       
  
  

## 4.2: Zoom in to check between January 1, 2021
## to April 30, 2021

interest_time %>%
  ggplot(aes(x = date, y = hits, color = keyword)) +
  geom_smooth(span = 0.1, se = FALSE) +
  facet_zoom(xlim =  c(as.POSIXct(Sys.Date()-60),as.POSIXct(Sys.Date()) )) +
  geom_point(color = "black") +
  geom_vline(xintercept = as.POSIXct(mx), color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Google Web search for 'Covid & Vaccine' in 2021 (Zooming last 60 days)",
       x = NULL, y = "Interest",
       caption = paste("Google Web Search Analysis :",Sys.Date()))

```

#Task Five: Related Queries Data
In this task, we will examine and 
explore related queries to Covid
```{r}

## 5.1: Return the related queries data
rel_queries <- trends$related_queries
rel_queries

library(tidyverse)
## 5.2: Create a bar plot for the queries 
## most related to covid

covid_search <- rel_queries %>%
  filter(grepl('covid', value) & related_queries == 'top' ) %>%
  mutate(subject = as.numeric(subject)) %>%
  arrange(desc(subject)) %>%
  mutate(order = c(1:nrow(covid_search)))

  

covid_search %>%
  ggplot(aes(x = reorder(value,-order), y = subject, fill = 'red', label = subject)) +
  geom_bar(stat = "identity", show.legend = F) +
  coord_flip() +
  theme_get() +
  labs(x = "Search Value", y = "Subject") +
  ggtitle("Google Queries most related to 'Covid' ") +
  theme(plot.title = element_text(colour = "red", size = 18),
        axis.title = element_text(colour = "blue")) +
  geom_text(size = 2, color = "blue", position = position_stack(vjust = 1.2))


  


  
  


```

#Task Six: Interest by Country Data
In this task, we will examine and 
explore the interest by country data
```{r}

## 6.1: Return the interest by country data
by_country <- wtrends$interest_by_country
by_country

## 6.2: Create a variable called "countries" 
## from the world data and interest by country data
countries <- spData::world %>%
  left_join(y = by_country, by = c("name_long" = "location"), keep = T)



## 6.3: Create a map of the interest by countries
tm_shape(countries) +
  tm_fill("hits",
          title = "Search Interest",
          legend.reverse = T,
          id = "name_long", 
          popup.vars = c(Name = "name_long", Search.Interest = "hits",
                       Population = "pop", Life.Expectancy = "lifeExp", 
                       GDP.per.capita = "gdpPercap"))

## 6.4: Different versions of the map
tmap_mode("view") #interactive maps

## OR

tmap_mode("plot") #plot-ready maps

```

#Task Seven: Interest by City Data
In this task, we will examine and
explore the interest by city data
```{r}

## 7.1: Return the interest by city data
interest_city <- wtrends$interest_by_city
interest_city

## 7.2: Return the top 3 cities with the highest hits
interest_city %>%
  arrange(desc(hits)) %>%
  head(3)

## 7.3: Return a table for the interest by city
datatable(interest_city)

## 7.4: Return the first 25 rows of the table
datatable(interest_city[c(1:25),])

```
